---
title: "Stats_prac_3_data2"
author: "Robby B"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(openxlsx)
library(ggplot2)
library(dplyr)
library(Rmisc)
library(knitr)
library(pastecs)
library(FSA)
library(nortest)
library(car)

```

# DATASET 2 f0.xlsx

### **Reading and preparing data**
```{r}
df <- read.xlsx("C:/Users/robby/OneDrive - University of Leeds/MSc Biodiversity and Conservation/BIodiv and Cons skills/Class 4/f0.xlsx")
kable(df)
dependent.var <- df[,1] #Column variabel dependent
independent.var <- df[,2]#Column variabel independent
satuan <- "(mg)"
df$Enzy <- factor(df$Enzy, levels = c(0,1,2), labels = c("control", "Enzyme1", "Enzyme2"))
```

### **Data descriptive summary**
```{r message=FALSE, warning=FALSE}
attach(df)
```

```{r}
df_summary <- data.frame(Summarize(Response ~ Enzy, digits = 2))
kable(df_summary, format = "pipe")
```

### **Homogeneity test**

#### Levene's test for homogeneity of variance.
*if p > 0.05 assumption of homogeneity fulfilled*
```{r}
levene <- suppressWarnings(leveneTest(Response ~ Enzy, data = df))
print(levene)
```

### **Normality test**
*if p > 0.05 assumption of normal distribution fulfilled*

#### Shapiro-Wilk Normality Test
```{r}
normalitas <- shapiro.test(Response)
print(normalitas)
```

#### Kolmogorov-Smirnov Tests
```{r}
uji_ks <- suppressWarnings(ks.test(Response,"pnorm",mean = mean(Response), 
                            sd = sd(Response)))
print(uji_ks)
```

#### Lilliefors (Kolmogorov-Smirnov) test for normality
Although the test statistic obtained from lillie.test(x) is the same as that obtained from ks.test(x, "pnorm", mean(x), sd(x)), it is not correct to use the p-value from the latter for the composite hypothesis of normality (mean and variance unknown), since the distribution of the test statistic is different when the parameters are estimated.

```{r}
lillie <- lillie.test(Response)
print(lillie)
```

### **Q-Q Plot**
A Q-Q plot is a scatterplot created by plotting two sets of quantiles against one another. If both sets of quantiles came from the same distribution, we should see the points forming a line that’s roughly straight.
```{r, fig.width=5.8, fig.height=4.5}
kol2 <- colnames(df)[1]
QQplot <- qqPlot(Response, id = F, ylab = kol2, main = "Quantile - Quantile Plot")
```

### **Box-and-whisker plot**
```{r fig.align='left', fig.width= 8, fig.height= 4}
box_whisker <- boxplot(Response ~ Enzy, 
                        xlab = paste(kol2,satuan), ylab = "", col = "orange", border = 
                        "brown", horizontal = TRUE, las = 1, par(mar = c(5,12,1,1), font.axis=2, font.lab=2))
```


##Non-Parametric analysis
This test is used when normality assumption of a dataset is not achieved.

```{r message=FALSE, warning=FALSE}
library(rcompanion)
library(multcompView)
library(DescTools)
library(FSA)
library(knitr)
```


### **Kruskal-Wallis Rank Sum Test**
```{r}
kruskalw <- kruskal.test(Response ~ Enzy)
print(kruskalw)
```

*Note: If Kruskal-Wallis test result is significant (p < 0.05) conitnue with post hoc test.

### **Post Hoc Tests**
*Test for several post hoc methods for comparison*

#### Pairwise Wilcoxon Rank Sum Tests
```{r error = T}
wilcox <- pairwise.wilcox.test(Response,Enzy,
                               p.adjust.method = "BH", exact = FALSE)
#p.adjust.method: Returns p-values adjusted using Benjamini & Hochberg (1995) method ("BH")
wilcox1 <- wilcox$p.value
wilcox2 <- fullPTable(wilcox1)
wilcox3 <- multcompLetters(wilcox2,compare = "<", threshold = 0.05, 
                           Letters = letters,reversed = T)
print(list(wilcox_pairwise = wilcox, wilcox3))
```


#### Dunn's Kruskal-Wallis Multiple Comparisons.
**An error indicates that there's no significant differences**
```{r error=T}
Dunn <- dunnTest(Response ~ Enzy, method = "bh")
Dunn1 <- Dunn$res
Dunn2 <- cldList(comparison = Dunn1$Comparison, p.value = Dunn1$P.adj, threshold = 0.05, reversed = T)
print(list(Dunn = Dunn, Dunn2))
```

#### Nemenyi Test
```{r}
Media <- as.factor(Enzy)
nemenyi <- NemenyiTest(x = Response, g = Enzy, dist = "tukey")
print(nemenyi)
```

**An error from cldList() indicate that there's no significant differences**
```{r error= T}
try(cldList(comparison = nemenyi$Comparison, p.value = nemenyi$P.adj, threshold = 0.05, reversed = T))
```

##Try with log transformation and then ANOVA

Log transformed
```{r}
df[,1] <- log10(df$Response)
```

**Use this analysis based on homogeneity and normality test results or a transformed  data**

```{r}
library(Rmisc)
library(agricolae)
library(ggplot2)
library(knitr)
```

```{r}
attach(df)
```

### **One-way ANOVA**
```{r}
model <- lm(Response ~ Enzy)
hasil_anova <- anova(model)
print(hasil_anova)
```

*Note: If ANOVA's result is significant (p < 0.05) continue with post hoc test

### **Summary linear model (Parameter estimates**
```{r}
summ <- summary(model)
print(summ)
```


### **Summarizes data**
*Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).*
```{r}
independent <- colnames(df)[2]
dependent <- colnames(df)[1]
dat <- summarySE(data = df,dependent,groupvars = independent,conf.interval = 0.95)
dat <- dat[order(dat[,3]),]
print(dat)
```

#Post Hoc Tests

### **Post Hoc: Tukey's HSD**
```{r}
independent <- colnames(df)[2]
dependent <- colnames(df)[1]
test_tukey <- HSD.test(model,colnames(df)[2], unbalanced = F)
print(test_tukey)
notasi <- test_tukey$groups[order(test_tukey$groups[,1]),]
notasi <- notasi[,2]
```

###**Bonferonni post hoc test (multiple comparisons)**

### **Summarizes data**
*Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).*
```{r}
dat <- summarySE(data = df,dependent,groupvars = independent,conf.interval = 0.95)
dat <- dat[order(dat[,3]),]
print(dat)
```
