---
title: "Stats_prac_2"
author: "Robby B"
date: "10/1/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Load Packages**
```{r}
library(haven)

```

**Read data**
```{r}
worms2010 <- read_sav("worms2010.sav")
worms2010 <- worms2010[-(67:75),]
#label sex
worms2010$sex <- factor(worms2010$sex, levels = c(1,2), labels = c("male", "female"))
attach(worms2010)
```

# A. Weight Data

**Graphs**
```{r}
#draw a boxplot of the data
boxplot(weight ~ sex, ylab = "Weight (g)")
boxplot.stats(weight)

#draw 2 histograms on separate rows
ggplot(worms2010, aes(x = weight)) +
  geom_histogram(aes(y = ..density..),fill = "white", colour = "black", bins = 20)+ 
  geom_density() +
  #creating normal distriution line
  stat_function(fun = dnorm, args = list(mean = mean(weight), sd = sd(weight)), colour = "red") +
  scale_x_continuous(name = "Weight (g)")  + 
  facet_grid(sex ~.) + ylab("Frequency")

#draw a bar charts with error bars. The default error bar is the 95% confidence interval
summary <- worms2010 %>% summarySE(measurevar = "weight", groupvars = c("sex"), 
                                        conf.interval = 0.95)
ggplot(summary, aes(x=sex, y=weight)) + 
  geom_pointrange(aes(ymin=weight-ci, ymax=weight+ci)) + 
  ylab("Weight (g)")

```

**Homogeneity test**
```{r}
#Levene test
levene <- suppressWarnings(leveneTest(weight ~ sex, data = worms2010, center = "mean"))
print(levene)
```

**normality test**
```{r}
mylist <- list(male = filter(worms2010, sex == "male")$weight, female = filter(worms2010, sex == "female")$weight)

#Shapiro-Wilk Normality Test
normalitas <- lapply(mylist, function (weight) shapiro.test(weight))
print(normalitas)

#Kolmogrov-Smirnov
uji_ks <- lapply(mylist, function(weight) suppressWarnings(ks.test(weight,"pnorm",mean = mean(weight), 
                            sd = sd(weight))))
print(uji_ks)

#Lilliefors (Kolmogorov-Smirnov) test for normality
lillie <- lapply(mylist, function(weight) {lillie.test(weight)})
print(lillie)

# Q-Q Plot
qqPlot(mylist[[1]], ylab = "Male Weight", id = F)
qqPlot(mylist[[2]], ylab = "Female Weight", id = F)

```

**Independent samples t tests**
```{r}
t.test(weight ~ sex, var.equal = T)
```

**Anova**
```{r}
model <- lm(weight ~ sex)
summary(model)
hasil_anova <- anova(model)
print(hasil_anova)
```

# B. Worms Data
**Graphs**
```{r}
#draw a boxplot of the data
boxplot(worms ~ sex, ylab = "Worms")
boxplot.stats(worms)

#draw 2 histograms on separate rows
ggplot(worms2010, aes(x = worms)) +
  geom_histogram(aes(y = ..density..),fill = "white", colour = "black", bins = 20)+ 
  geom_density() +
  #creating normal distriution line
  stat_function(fun = dnorm, args = list(mean = mean(worms), sd = sd(worms)), colour = "red") +
  scale_x_continuous(name = "worms")  + 
  facet_grid(sex ~.) + ylab("Frequency")

```

**Homogeneity test**
```{r}
#Levene test
levene <- suppressWarnings(leveneTest(worms ~ sex, data = worms2010, center = "mean"))
print(levene)
```

**normality test**
```{r}
mylist <- list(male = filter(worms2010, sex == "male")$worms, female = filter(worms2010, sex == "female")$worms)

#Shapiro-Wilk Normality Test
normalitas <- lapply(mylist, function (worms) shapiro.test(worms))
print(normalitas)

#Kolmogrov-Smirnov
uji_ks <- lapply(mylist, function(worms) suppressWarnings(ks.test(worms,"pnorm",mean = mean(worms), 
                            sd = sd(worms))))
print(uji_ks)

#Lilliefors (Kolmogorov-Smirnov) test for normality
lillie <- lapply(mylist, function(worms) {lillie.test(worms)})
print(lillie)

# Q-Q Plot
qqPlot(mylist[[1]], ylab = "Male Worm", id = F)
qqPlot(mylist[[2]], ylab = "Female Worm", id = F)

```

**Non parametric tests**
```{r}
#The Mann-Whitney test
wilcox.test(worms ~ sex)

#Kruskal Wallis test --> same with the spss result using mann whitney
kruskalw <- kruskal.test(worms ~ sex)
print(kruskalw)
```

#C. Paired data

**Read data**
```{r}
paired_data <- read_sav("pairs(1).sav")
attach(paired_data)
```

**t test**
```{r}
#paired t test
t.test(pre, post, paired = T)
#one sample t test on the difference between pre and post
t.test(difference)
```
both approaches have the same result

**check normality**
```{r}
shapiro.test(difference)
lillie.test(difference)
```
The difference of pre and post is not normally distributed, thus, t test result is not valid

**Non parametric test**
```{r}
#Wilcoxon Rank-Sum test 
wilcox.test(difference)
```
