---
title: "SDM_logit_future"
output: word_document
date: '2022-06-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library("sp")
library("raster")
library("maptools")
library("rgdal")
library("dismo")
library("sf")
library("terra")
library("geodata")
``` 

[1] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp126_2021-2040_masked.tif
[2] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp126_2041-2060_masked.tif
[3] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp370_2021-2040_masked.tif
[4] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp370_2041-2060_masked.tif
# Load environmental variables
```{r}
predictors <- list.files("~/My R/Advanced_Statistics/summerproject/species_modelling/future",pattern = "*.tif$", full.names = TRUE)
predictors <- brick(predictors[1])
names(predictors) <- c("bio1", "bio2", "bio3", "bio4", "bio5", "bio6", "bio7", "bio8",
                               "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15",
                               "bio16", "bio17", "bio18", "bio19")
```

# Load observation data and Aceh Boundary
```{r}
obs.data <- readOGR("~/My R/Advanced_Statistics/summerproject/species_modelling/all_dataset_new.shp")
names(obs.data)[2]<- "latitude"
names(obs.data)[3] <- "longitude"

aceh <- st_read("~/My R/Advanced_Statistics/summerproject/BATAS_PROVINSI_ACEH.shp")
colnames(obs.data@coords)[1] <- "longitude"
colnames(obs.data@coords)[2] <- "latitude"

```

# Set geographic extent
```{r}
#Determine geographic extent of our data
ext <- extent(predictors)

```

# Extract environmental values based on presence data
```{r}
#presvals <- extract(predictors, obs.data@coords)

```

# Create random points

```{r}
# Use the bioclim data files for sampling resolution
bil.files <- list.files(path = "~/My R/Advanced_Statistics/summerproject/species_modelling/historical/20182019", 
                          pattern = "*.tif$", 
                          full.names = TRUE)

# We only need one file, so use the first one in the list of .bil files
mask <- raster(bil.files[1])

# Set the seed for the random-number generator to ensure results are similar
set.seed(20210707)

# Randomly sample points (same number as our observed points)
backgr <- randomPoints(mask = predictors,     # Provides resolution of sampling points
                   n = nrow(obs.data),      # Number of random points
                   ext = ext, # Spatially restricts sampling
                   extf = 1.25)             # Expands sampling a little bit

colnames(backgr) = c('longitude', 'latitude')
```

# Plot the pseudo-absence points and observation data
```{r, fig.dim= c(6,8)}
plot(aceh$geometry,
     axes = TRUE, 
     col = "grey95",
     main = "Presence and pseudo-absence points")

# Add the background points
points(backgr, col = "grey30", pch = 1, cex = 0.75)

# Add the observations
points(obs.data@coords, 
       col = "olivedrab", 
       pch = 20, 
       cex = 0.75)

box()

```


# Extract environmental values based on background data
```{r}
#absvals <- extract(predictors, backgr)
```

# Combine extracted presence and background values into a single data.frame in which the first column (variable ‘pb’) indicates whether this is a presence or a background point. ‘Forest_status’ is categorical variable (called a ‘factor’ in R) and it is important to explicitly define it that way, so that it won’t be treated like any other numerical variable. 
```{r}
#pb <- c(rep(1, nrow(presvals)), rep(0, nrow(absvals)))
#sdmdata <- data.frame(cbind(pb, rbind(presvals, absvals)))
#sdmdata[,'Forest_status'] <- as.factor(sdmdata[,'Forest_status'])
```

#create train and test data for presence
```{r}
# Arbitrarily assign group 1 as the testing data group
testing.group <- 1

# Create vector of group memberships
set.seed(0)
group <- kfold(x = obs.data@coords, k = 5) # kfold is in dismo package

head(group)
# Should see even representation in each group
table(group)

```

```{r}
# Separate observations into training and testing groups
pres_train <- obs.data@coords[group != testing.group, ]
pres_test <- obs.data@coords[group == testing.group, ]
```

#create train and test data for absence
```{r}
# Repeat the process for pseudo-absence points
set.seed(10)
group.background <- kfold(x = backgr, k = 5)
backg_train <- backgr[group.background != testing.group, ]
backg_test <- backgr[group.background == testing.group, ]

```

#Plot
```{r}
r <- raster(predictors, 1)
plot(!is.na(r), col=c('white', 'light grey'), legend=FALSE)
plot(ext, add=TRUE, col='red', lwd=2)
points(backg_train, pch='-', cex=0.5, col='yellow')
points(backg_test, pch='-',  cex=0.5, col='black')
points(pres_train, pch= '+', col='green')
points(pres_test, pch='+', col='blue')
```

We now have a set of predictor variables (rasters) and occurrence points. The next step is to extract the values of the predictors at the locations of the points. (This step can be skipped for the modeling methods that are implemented in the dismo package). This is a very straightforward thing to do using the ‘extract’ function from the raster package. In the example below we use that function first for the Bradypus occurrence points, then for random background points.

#Extract environmental values based on presence and absence train coordinates
```{r}
train <- rbind(pres_train, backg_train)
pb_train <- c(rep(1, nrow(pres_train)), rep(0, nrow(backg_train)))
envtrain <- extract(predictors, train)
envtrain <- data.frame( cbind(pa=pb_train, envtrain) )
head(envtrain)
```

#Extract environmental values based on presence and absence test coordinates
```{r}
testpres <- data.frame( extract(predictors, pres_test) )
testbackg <- data.frame( extract(predictors, backg_test) )
```

#Generalized Linear Models
```{r}
# logistic regression:
gm1 <- glm(pa ~ bio1+bio2+bio3+bio4+bio5+bio6+bio7+bio8+bio9+bio10+bio11+bio12+bio13+bio14+bio15+bio16+bio17+bio18+bio19,
            family = binomial(link = "logit"), data=envtrain)
summary(gm1)
coef(gm1)

# Gaussian
gm2 <- glm(pa ~ bio1+bio2+bio3+bio4+bio5+bio6+bio7+bio8+bio9+bio10+bio11+bio12+bio13+bio14+bio15+bio16+bio17+bio18+bio19,
            family = gaussian(link = "identity"), data=envtrain)
summary(gm2)
coef(gm2)

# Poisson
gm3 <- glm(pa ~ bio1+bio2+bio3+bio4+bio5+bio6+bio7+bio8+bio9+bio10+bio11+bio12+bio13+bio14+bio15+bio16+bio17+bio18+bio19,
            family = poisson(link = "log"), data=envtrain)
summary(gm3)
coef(gm3)
```

# Evaluate model and indicate the best model with AIC and AUC scores
choose the highest AUC and the lowest AIC
```{r}
ge1 <- evaluate(testpres, testbackg, gm1)
ge2 <- evaluate(testpres, testbackg, gm2)
ge3 <- evaluate(testpres, testbackg, gm3)
print(c(ge1,ge2,ge3))
print (c(gm1[["aic"]],gm2[["aic"]],gm3[["aic"]]))
```

#Predict the model
```{r}
pg <- predict(predictors, gm2, ext=ext)
```

```{r}
par(mfrow=c(1,2))
plot(pg, main='GLM, raw values')
plot(aceh$geometry, add=TRUE, border='dark grey')
```

#generate threshold and plot it
```{r}
stat_used <- "spec_sens"
tr <- threshold(ge2, stat_used)

plot(pg > tr, main='presence/absence')
plot(aceh$geometry, add=TRUE, border='dark grey')
#points(pres_train, pch='+')
#points(backg_train, pch='-', cex=0.25)

```

# Generate a map with threshold
```{r}
predict_threshold <- pg > tr
rm <- c(0, NA)
reclassmatrix  <- matrix(rm, ncol = 2)
predict_threshold <- reclassify(predict_threshold, reclassmatrix)

```

# Write raster ONCE
```{r}

raster::writeRaster(pg, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/future/regression/future_predicted_GLM.tif"), overwrite = TRUE)
```

# Write raster depends on threshold
```{r}
threshold_used <- paste(tr, stat_used)
date_now <- format(Sys.time(), "%Y-%m-%d %H-%M-%S")

raster::writeRaster(predict_threshold, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/future/regression/future_predicted_GLM_",stat_used,".tif"), overwrite = TRUE)

write.table(threshold_used, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/future/regression/future_threshold_",stat_used,"_",date_now,".txt"))

```