---
title: "Species_modelling historical"
subs
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library("sp")
library("raster")
library("maptools")
library("rgdal")
library("dismo")
library("sf")
library("terra")
library("geodata")
``` 

# Load environmental variables
```{r}

bioclim.data <- list.files("~/My R/Advanced_Statistics/summerproject/species_modelling/historical/20182019",pattern = "*.tif$", full.names = TRUE)
bioclim.data <- lapply(bioclim.data, raster)
bioclim.data <- stack(bioclim.data)
bioclim.data$Tree_cover  <- extractByMask(bioclim.data$Tree_cover, bioclim.data$Distance_to_Crops, spatial = T)
matr <- c(100, bioclim.data@layers[[8]]@data@max, 0)
rec_matrix <- matrix(matr, ncol = 3)
bioclim.data$Tree_cover <- reclassify(bioclim.data$Tree_cover, rec_matrix)

bioclim.data <- dropLayer(bioclim.data, "Slope")
```

# Load observation data and Aceh Boundary
```{r}
obs.data <- readOGR("~/My R/Advanced_Statistics/summerproject/species_modelling/all_dataset_new.shp")
names(obs.data)[2]<- "latitude"
names(obs.data)[3] <- "longitude"

aceh <- st_read("~/My R/Advanced_Statistics/summerproject/BATAS_PROVINSI_ACEH.shp")
colnames(obs.data@coords)[1] <- "longitude"
colnames(obs.data@coords)[2] <- "latitude"

```

# Set geographic extent
```{r}
#Determine geographic extent of our data
geographic.extent <- extent(bioclim.data)

```

#plot observation data and Aceh boundary
```{r, fig.dim= c(6,8)}
# Plot the base map
plot(aceh$geometry, 
     axes = TRUE, 
     col = "grey95")

# Add the points for individual observation
points(obs.data@coords, 
       col = "olivedrab")
# And draw a little box around the graph
box()


```

# Build the distribution model
```{r}

# Build species distribution model
bc.model <- bioclim(x = bioclim.data, p = obs.data@coords)
response(bc.model)

```

# Predict the distribution model
```{r}
# Predict presence from model
predict.presence.historical <- dismo::predict(object = bc.model, 
                                   x = bioclim.data,
                                   ext=  geographic.extent)

```


```{r, fig.dim= c(6,8)}
# Plot Boundaries
plot(aceh$geometry, 
     axes = TRUE, 
     col = "grey95")

# Add model probabilities
plot(predict.presence.historical, add = TRUE)

#points(obs.data@coords, 
       #col = "olivedrab")
box()

```

#The pseudo-absence point

```{r}
# Use the bioclim data files for sampling resolution
bil.files <- list.files(path = "~/My R/Advanced_Statistics/summerproject/species_modelling/historical/20182019", 
                          pattern = "*.tif$", 
                          full.names = TRUE)

# We only need one file, so use the first one in the list of .bil files
mask <- raster(bil.files[1])

# Set the seed for the random-number generator to ensure results are similar
set.seed(20210707)

# Randomly sample points (same number as our observed points)
background <- randomPoints(mask = mask,     # Provides resolution of sampling points
                   n = nrow(obs.data),      # Number of random points
                   ext = geographic.extent, # Spatially restricts sampling
                   extf = 1.25)             # Expands sampling a little bit

```

# Plot the pseudo-absence points and observation data
```{r, fig.dim= c(6,8)}
plot(aceh$geometry,
     axes = TRUE, 
     col = "grey95",
     main = "Presence and pseudo-absence points")

# Add the background points
points(background, col = "grey30", pch = 1, cex = 0.75)

# Add the observations
points(obs.data@coords, 
       col = "olivedrab", 
       pch = 20, 
       cex = 0.75)

box()

```

#create train and test data for pseudo-absence
```{r}
# Arbitrarily assign group 1 as the testing data group
testing.group <- 1

# Create vector of group memberships
group.presence <- kfold(x = obs.data@coords, k = 5) # kfold is in dismo package
```

```{r}
head(group.presence)
# Should see even representation in each group
table(group.presence)

```

```{r}
# Separate observations into training and testing groups
presence.train <- obs.data[group.presence != testing.group, ]
presence.test <- obs.data[group.presence == testing.group, ]

# Repeat the process for pseudo-absence points
group.background <- kfold(x = background, k = 5)
background.train <- background[group.background != testing.group, ]
background.test <- background[group.background == testing.group, ]

```

#Training and testing the model and predict

```{r}
# Build a model using training data
bc.model <- bioclim(x = bioclim.data, p =presence.train@coords)

# Predict presence from model (same as previously, but with the update model)
predict.presence.pseudo <- dismo::predict(object = bc.model, 
                                   x = bioclim.data, 
                                   ext = geographic.extent)
response(bc.model)

```

# Generate Threshold
```{r}
# Use testing data for model evaluation
bc.eval <- evaluate(p = presence.test@coords,   # The presence testing data
                    a = background.test, # The absence testing data
                    model = bc.model,    # The model we are evaluating
                    x = bioclim.data)    # Climatic variables for use by model

# Determine minimum threshold for "presence"
stat_used <- "spec_sens"
bc.threshold <- threshold(x = bc.eval, stat = stat_used)

```

# Generate a map with threshold
```{r}
reclass_predict.presence.pseudo <- predict.presence.pseudo > bc.threshold
rm <- c(0, NA)
reclassmatrix  <- matrix(rm, ncol = 2)
reclass_predict.presence.pseudo <- reclassify(reclass_predict.presence.pseudo, reclassmatrix)

```

#Plot the predicted distribution
```{r, fig.dim= c(6,8)}
# Plot base map
plot(aceh$geometry, 
     axes = TRUE, 
     col = "grey95")

# Add model probabilities
plot(predict.presence.pseudo, add = TRUE)

# Add original observations
#points(obs.data@coords, col = "black", pch = "+", cex = 0.75)
box()
```

#plot the map with threshold
```{r, fig.dim= c(6,8)}
# Plot base map
plot(aceh$geometry,
     axes = TRUE, 
     col = "grey95")

# Only plot areas where probability of occurrence is greater than the threshold
plot(predict.presence.pseudo > bc.threshold, 
     add = TRUE, 
     legend = FALSE, 
     col = c(NA, "olivedrab"))

# And add those observations
#points(obs.data@coords, 
       #col = "black",
       #pch = "+", 
       #cex = 0.75)

box()

```

# Write raster ONCE
```{r}
raster::writeRaster(predict.presence.historical, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/bioclim/Historical_presence_only.tif"), overwrite = TRUE)

raster::writeRaster(predict.presence.pseudo, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/bioclim/Historical_presence_absence.tif"), overwrite = TRUE)
```

# Write raster depends on threshold
```{r}
threshold_used <- paste(bc.threshold, stat_used)
date_now <- format(Sys.time(), "%Y-%m-%d %H-%M-%S")

raster::writeRaster(reclass_predict.presence.pseudo, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/bioclim/Historical_presence_absence_",stat_used,".tif"), overwrite = TRUE)

write.table(threshold_used, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/bioclim/threshold_",stat_used,"_",date_now,".txt"))

```
