---
title: "SDM Maxent Future"
output: word_document
date: '2022-06-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library("sp")
library("raster")
library("maptools")
library("rgdal")
library("dismo")
library("sf")
library("terra")
library("geodata")
library("rJava")
``` 

[1] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp126_2021-2040_masked.tif
[2] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp126_2041-2060_masked.tif
[3] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp370_2021-2040_masked.tif
[4] wc2.1_2.5m_bioc_CNRM-CM6-1_ssp370_2041-2060_masked.tif
# Load environmental variables
```{r}
predictors <- list.files("~/My R/Advanced_Statistics/summerproject/species_modelling/future",pattern = "*.tif$", full.names = TRUE)
predictors <- brick(predictors[1])
names(predictors) <- c("bio1", "bio2", "bio3", "bio4", "bio5", "bio6", "bio7", "bio8",
                               "bio9", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15",
                               "bio16", "bio17", "bio18", "bio19")
```

# Load observation data and Aceh Boundary
```{r}
obs.data <- readOGR("~/My R/Advanced_Statistics/summerproject/species_modelling/all_dataset_new.shp")
names(obs.data)[2]<- "latitude"
names(obs.data)[3] <- "longitude"

aceh <- st_read("~/My R/Advanced_Statistics/summerproject/BATAS_PROVINSI_ACEH.shp")
colnames(obs.data@coords)[1] <- "longitude"
colnames(obs.data@coords)[2] <- "latitude"

```

# Set geographic extent
```{r}
#Determine geographic extent of our data
ext <- extent(predictors)

```


# Create random points

```{r}
# Use the bioclim data files for sampling resolution
bil.files <- list.files(path = "~/My R/Advanced_Statistics/summerproject/species_modelling/historical/20182019", 
                          pattern = "*.tif$", 
                          full.names = TRUE)

# We only need one file, so use the first one in the list of .bil files
mask <- raster(bil.files[1])

# Set the seed for the random-number generator to ensure results are similar
set.seed(20210707)

# Randomly sample points (same number as our observed points)
backgr <- randomPoints(mask = predictors,     # Provides resolution of sampling points
                   n = nrow(obs.data),      # Number of random points
                   ext = ext, # Spatially restricts sampling
                   extf = 1.25)             # Expands sampling a little bit

colnames(backgr) = c('longitude', 'latitude')
```

# Plot the pseudo-absence points and observation data
```{r, fig.dim= c(6,8)}
plot(aceh$geometry,
     axes = TRUE, 
     col = "grey95",
     main = "Presence and pseudo-absence points")

# Add the background points
points(backgr, col = "grey30", pch = 1, cex = 0.75)

# Add the observations
points(obs.data@coords, 
       col = "olivedrab", 
       pch = 20, 
       cex = 0.75)

box()

```


#create train and test data for presence
```{r}
# Arbitrarily assign group 1 as the testing data group
testing.group <- 1

# Create vector of group memberships
set.seed(0)
group <- kfold(x = obs.data@coords, k = 5) # kfold is in dismo package

head(group)
# Should see even representation in each group
table(group)

```

```{r}
# Separate observations into training and testing groups
pres_train <- obs.data@coords[group != testing.group, ]
pres_test <- obs.data@coords[group == testing.group, ]
```

#create train and test data for absence
```{r}
# Repeat the process for pseudo-absence points
set.seed(10)
group.background <- kfold(x = backgr, k = 5)
backg_train <- backgr[group.background != testing.group, ]
backg_test <- backgr[group.background == testing.group, ]

```

#Plot
```{r}
r <- raster(predictors, 1)
plot(!is.na(r), col=c('white', 'light grey'), legend=FALSE)
plot(ext, add=TRUE, col='red', lwd=2)
points(backg_train, pch='-', cex=0.5, col='yellow')
points(backg_test, pch='-',  cex=0.5, col='black')
points(pres_train, pch= '+', col='green')
points(pres_test, pch='+', col='blue')
```


#Maxent Models
```{r}
maxent()
xm <- maxent(predictors, pres_train,backg_train)
plot(xm)
response(xm)

```

# Evaluate model and indicate the best model with AIC and AUC scores
```{r}
e <- evaluate(pres_test, backg_test, xm, predictors)
e

```

#Predict the model
```{r}
px <- predict(predictors, xm, ext=ext, progress='')
```

```{r}
par(mfrow=c(1,2))
plot(px, main='Maxent, raw values')
plot(aceh$geometry, add=TRUE, border='dark grey')
```

#generate threshold and plot it
```{r}
stat_used <- "sensitivity"
tr <- threshold(e, stat_used)
plot(px > tr, main='presence/absence')
plot(aceh$geometry, add=TRUE, border='dark grey')
#points(pres_train, pch='+')

```

# Generate a map with threshold
```{r}
predict_threshold <- px > tr
rm <- c(0, NA)
reclassmatrix  <- matrix(rm, ncol = 2)
predict_threshold <- reclassify(predict_threshold, reclassmatrix)

```

# Write raster ONCE
```{r}

raster::writeRaster(px, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/future/maxent/future_predicted_maxent.tif"), overwrite = TRUE)
```

# Write raster depends on threshold
```{r}
threshold_used <- paste(tr, stat_used)
date_now <- format(Sys.time(), "%Y-%m-%d %H-%M-%S")

raster::writeRaster(predict_threshold, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/future/maxent/future_predicted_maxent_",stat_used,".tif"), overwrite = TRUE)

write.table(threshold_used, paste0("~/My R/Advanced_Statistics/summerproject/species_modelling/results/future/maxent/future_threshold_",stat_used,"_",date_now,".txt"))

```