---
title: "Homogeneity and normality"
subtitle: "Population dynamics part 2"
author: "Robby"
date: "11/30/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(FSA)
library(car)
library(nortest)
library(knitr)
library(haven)

```


### **Reading and preparing data**
```{r}
setwd("~/My R/Advanced_Statistics/pop_dynamics")
df <- read_sav("data.sav")
kable(df)
dependent.var <- df[,3] #Column variabel dependent
independent.var <- df[,1]#Column variabel independent
df$Fishery <- factor(df$Fishery, levels = c(0,1), labels = c("without", "with"))
#satuan <- "(cm)"
```

### **Data descriptive summary**
```{r message=FALSE, warning=FALSE}
attach(df)
```

```{r}
df_summary <- data.frame(Summarize(logabundance ~ Fishery, digits = 2))
kable(df_summary, format = "pipe")
```

### **Homogeneity test**

#### Levene's test for homogeneity of variance.
*if p > 0.05 assumption of homogeneity fulfilled*
```{r}
levene <- suppressWarnings(leveneTest(logabundance ~ Fishery, data = df))
print(levene)
```

### **Normality test**
*if p > 0.05 assumption of normal distribution fulfilled*

#### Shapiro-Wilk Normality Test
```{r}
logabundance_with <- subset(logabundance, Fishery == "with")
logabundance_without <- subset(logabundance, Fishery == "without")
normalitas_with <- shapiro.test(logabundance_with)
normalitas_without <- shapiro.test(logabundance_without)
print(normalitas_with)
print(normalitas_without)
```

#### Kolmogorov-Smirnov Tests
```{r}
uji_ks <- suppressWarnings(ks.test(logabundance,"pnorm",mean = mean(logabundance), 
                            sd = sd(logabundance)))
print(uji_ks)
```

#### Lilliefors (Kolmogorov-Smirnov) test for normality
Although the test statistic obtained from lillie.test(x) is the same as that obtained from ks.test(x, "pnorm", mean(x), sd(x)), it is not correct to use the p-value from the latter for the composite hypothesis of normality (mean and variance unknown), since the distribution of the test statistic is different when the parameters are estimated.

```{r}
lillie <- lillie.test(logabundance)
print(lillie)
```

### **Q-Q Plot**
A Q-Q plot is a scatterplot created by plotting two sets of quantiles against one another. If both sets of quantiles came from the same distribution, we should see the points forming a line that’s roughly straight.
```{r, fig.width=5.8, fig.height=4.5}
kol2 <- colnames(df)[2]
QQplot <- qqPlot(logabundance, id = F, ylab = kol2, main = "Quantile - Quantile Plot")
```

### **Box-and-whisker plot**
```{r fig.align='left', fig.width= 8, fig.height= 4}
box_whisker <- boxplot(logabundance ~ Fishery, 
                        xlab = "Log Abundance", ylab = "", col = "orange", border = 
                        "brown", horizontal = TRUE, las = 1, par(mar = c(5,12,1,1), font.axis=2, font.lab=2))
```


